---
title: "sim"
author: "lwa19"
date: "2022-09-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

For each # of dice we can reroll, simulate all possible outcomes and their score outcome

## Scoring scheme: 

Input the dice roll results, give the possible dice rolls and the total number of points

```{r score}
# input "rolls" as a vector
score <- function(rolls) {
  nrolls <- length(rolls)
  ord_rl <- sort(rolls)
  df = as.data.frame(table(ord_rl))
  uniq_rolls = length(df$ord_rl)
  
  # 6 dice
  if (nrolls == 6) {
    ret_string = character()
    curr = 0
    if (uniq_rolls == 6) {
      #print('full run: 2500')
      return(2500)
    } else if ((uniq_rolls == 3) & (all(df$Freq == 2))) {
      #print('three pairs: 1500')
      return(1500)
    } else if (uniq_rolls == 1) {
      #print('six of a kind: 3000')
      return(3000)
    } else if ((uniq_rolls == 2) & (any(df$Freq) == 5)) {
      curr = 2000
      ret_string = "five of a kind"
      
      # check for 1/5
      single_id = which(df$Freq != 5)
      single_val = df$ord_rl[single_id]

      if (single_val == 1) {
        curr = curr + 100
        ret_string = paste0(ret_string, '; ones')
      } else if (single_val == 5) {
        curr = curr + 50
        ret_string = paste0(ret_string, '; fives')
      } 
      return(curr)
    } else if (any(df$Freq == 4)) {
      curr = 1000
      ret_string = "four of a kind"
      
      # check for 1s/5s
      ones = which(df$ord_rl == 1)
      if (length(ones) == 1) {
        curr = curr + 100 * df$Freq[ones]
        ret_string = paste0(ret_string, '; ones')
      }
      fives = which(df$ord_rl == 5)
      if (length(fives) == 1) {
        curr = curr + 50 * df$Freq[fives]
        ret_string = paste0(ret_string, '; fives')
      }
    } else if (any(df$Freq == 3)) {
      triple_id = which(df$Freq == 3)
      for (id in triple_id) {
        if (df$ord_rl[id] == 1) {
          curr = curr + 1000
          ret_string = paste0(ret_string, 'triple 1')
        } else {
          curr = curr + ord_rl[id] * 100
          ret_string = paste0(ret_string, 'triple', df$ord_rl[id])
        }
      }
      
      # check for 1s/5s
      rem = df$ord_rl[-triple_id]
      freq_rem = df$Freq[-triple_id]
      ones = which(rem == 1)
      if (length(ones) == 1) {
        curr = curr + 100 * freq_rem[ones]
        ret_string = paste0(ret_string, '; ones')
      }
      fives = which(rem == 5)
      if (length(fives) == 1) {
        curr = curr + 50 * freq_rem[fives]
        ret_string = paste0(ret_string, '; fives')
      }
    } else if (any(df$ord_rl == 1) | any(df$ord_rl == 5)) {
      # check ones
      ones_id = which(df$ord_rl == 1)
      if (length(ones_id) > 0) {
        curr = curr + df$Freq[ones_id] * 100
        ret_string = paste0(ret_string, 'ones')
      }
       
      # check fives
      fives_id = which(df$ord_rl == 5)
      if (length(fives_id) > 0) {
        curr = curr + df$Freq[fives_id] * 50
        ret_string = paste0(ret_string, 'fives')
      }
    }
    #print(paste(ret_string, ':', curr))
    return(curr)
# five dice: 
  } else if (nrolls == 5) {
    curr = 0
    ret_string = character()
    
    # check five of a kind
    if (uniq_rolls == 1) {
      curr = 2000
      ret_string = "five of a kind"
      
    # check 4 of a kind: 
    } else if ((uniq_rolls == 2) & (any(df$Freq) == 4)) {
      curr = 1000
      ret_string = "four of a kind"
      
      # check for 1/5
      single_id = which(df$Freq != 5)
      single_val = df$ord_rl[single_id]

      if (single_val == 1) {
        curr = curr + 100
        ret_string = paste0(ret_string, '; ones')
      } else if (single_val == 5) {
        curr = curr + 50
        ret_string = paste0(ret_string, '; fives')
      } 
      
    # check triple: 
    } else if (any(df$Freq == 3)) {
      triple_id = which(df$Freq == 3)
      curr = 0
      ret_string = character()
      
      if (df$ord_rl[triple_id] == 1) {
        curr = curr + 1000
        ret_string = paste0(ret_string, 'triple 1')
      } else {
        curr = curr + ord_rl[triple_id] * 100
        ret_string = paste0(ret_string, 'triple', df$ord_rl[triple_id])
      }
      
      # check for 1s/5s
      rem = df$ord_rl[-triple_id]
      freq_rem = df$Freq[-triple_id]
      ones = which(rem == 1)
      if (length(ones) == 1) {
        curr = curr + 100 * freq_rem[ones]
        ret_string = paste0(ret_string, '; ones')
      }
      fives = which(rem == 5)
      if (length(fives) == 1) {
        curr = curr + 50 * freq_rem[fives]
        ret_string = paste0(ret_string, '; fives')
      }
    
    # check 1s/5s
    } else if (any(df$ord_rl == 1) | any(df$ord_rl == 5)) {
      # check ones
      ones_id = which(df$ord_rl == 1)
      if (length(ones_id) > 0) {
        curr = curr + df$Freq[ones_id] * 100
        ret_string = paste0(ret_string, 'ones')
      }
       
      # check fives
      fives_id = which(df$ord_rl == 5)
      if (length(fives_id) > 0) {
        curr = curr + df$Freq[fives_id] * 50
        ret_string = paste0(ret_string, 'fives')
      }
    }
    #print(paste(ret_string, ':', curr))
    return(curr)
    
# 4 dice: 
  } else if (nrolls == 4) {
    curr = 0
    ret_string = character()
    
    # check four of a kind
    if (uniq_rolls == 1) {
      curr = 1000
      ret_string = "four of a kind"
      
    # check triple: 
    } else if (any(df$Freq == 3)) {
      triple_id = which(df$Freq == 3)
      curr = 0
      ret_string = character()
      
      if (df$ord_rl[triple_id] == 1) {
        curr = curr + 1000
        ret_string = paste0(ret_string, 'triple 1')
      } else {
        curr = curr + ord_rl[triple_id] * 100
        ret_string = paste0(ret_string, 'triple', df$ord_rl[triple_id])
      }
      
      # check for 1s/5s
      rem = df$ord_rl[-triple_id]
      freq_rem = df$Freq[-triple_id]
      ones = which(rem == 1)
      if (length(ones) == 1) {
        curr = curr + 100 * freq_rem[ones]
        ret_string = paste0(ret_string, '; ones')
      }
      fives = which(rem == 5)
      if (length(fives) == 1) {
        curr = curr + 50 * freq_rem[fives]
        ret_string = paste0(ret_string, '; fives')
      }
    
    # check 1s/5s: 
    } else if (any(df$ord_rl == 1) | any(df$ord_rl == 5)) {
      # check ones
      ones_id = which(df$ord_rl == 1)
      if (length(ones_id) > 0) {
        curr = curr + df$Freq[ones_id] * 100
        ret_string = paste0(ret_string, 'ones')
      }
       
      # check fives
      fives_id = which(df$ord_rl == 5)
      if (length(fives_id) > 0) {
        curr = curr + df$Freq[fives_id] * 50
        ret_string = paste0(ret_string, 'fives')
      }
    }
    #print(paste(ret_string, ':', curr))
    return(curr)

# 3 dice
  } else if (nrolls == 3) {
    curr = 0
    ret_string = character()
    
    # check triple
    if (uniq_rolls == 1) {
      triple_id = which(df$Freq == 3)
      curr = 0
      ret_string = character()
      
      if (df$ord_rl[triple_id] == 1) {
        curr = curr + 1000
        ret_string = paste0(ret_string, 'triple 1')
      } else {
        curr = curr + ord_rl[triple_id] * 100
        ret_string = paste0(ret_string, 'triple', df$ord_rl[triple_id])
      }
    
    # check 1s/5s
    } else if (any(df$ord_rl == 1) | any(df$ord_rl == 5)) {
      # check ones
      ones_id = which(df$ord_rl == 1)
      if (length(ones_id) > 0) {
        curr = curr + df$Freq[ones_id] * 100
        ret_string = paste0(ret_string, 'ones')
      }
       
      # check fives
      fives_id = which(df$ord_rl == 5)
      if (length(fives_id) > 0) {
        curr = curr + df$Freq[fives_id] * 50
        ret_string = paste0(ret_string, 'fives')
      }
    }
    #print(paste(ret_string, ':', curr))
    return(curr)

# 2 or 1 dice: 
  } else {
    curr = 0
    ret_string = character()
    
    # check ones
      ones_id = which(df$ord_rl == 1)
      if (length(ones_id) > 0) {
        curr = curr + df$Freq[ones_id] * 100
        ret_string = paste0(ret_string, 'ones')
      }
       
      # check fives
      fives_id = which(df$ord_rl == 5)
      if (length(fives_id) > 0) {
        curr = curr + df$Freq[fives_id] * 50
        ret_string = paste0(ret_string, 'fives')
      }
    
    #print(paste(ret_string, ':', curr))
    return(curr)
  }
}
```

### Examples: 

```{r}
# 6 dice: 
score(c(1,2,3,4,5,6)) # [1] "full run: 2500"

rolls = c(3, 3, 5, 2, 3, 4)
score(rolls) # "triple3; fives: 350"

rolls = c(6, 1, 5, 6, 1, 5)
score(rolls) # "three pairs: 1500"

rolls = c(5, 4, 5, 6, 2, 4)
score(rolls) # "fives : 100"

rolls = c(1, 1, 1, 3, 5, 5)
score(rolls) # "triple 1; fives: 1100"

rolls = c(2, 1, 4, 6, 6, 1)
score(rolls) # "ones : 200"

rolls = c(2, 2, 6, 1, 1, 6)
score(rolls) # "three pairs: 1500"

rolls = c(1, 3, 5, 4, 1, 3)
score(rolls) # "onesfives : 250"

# 5 dice: 
rolls = c(3, 6, 6, 1, 2)
score(rolls) # "ones : 100"
  
# 4 dice: 
rolls = c(2, 4, 5, 6)
score(rolls) # "fives : 50"

rolls = c(1, 4, 2, 5)
score(rolls) # "onesfives : 150"

# 3 dice: 
rolls = c(5, 2, 2)
score(rolls) # "fives : 50"

rolls = c(4, 2, 2)
score(rolls) # " : 0"

# 2 dice
rolls = c(1, 2)
score(rolls) # "ones : 100"

rolls = c(5, 4)
score(rolls) # "fives : 50"

# 1 die
rolls = 3 
score(rolls) # " : 0" 

rolls = 1
score(rolls) # "ones : 100"
```

## Get scoring average for each # of dice rolls: 

```{r}
dice = 1:6

for (d in dice) {
  nsim = 10000
  total_score = 0
  for (i in 1:nsim) {
    rolls = sample(6, d, replace = T)
    sc = score(rolls)
    if (is.na(sc)) {
      break(paste('rolls scored as NA:', rolls))
    }
    total_score = total_score + score(rolls)
  }
  print(paste(d, 'dice roll has average score:', total_score / nsim))
}
```

Results (an estimate): 

- one die: 25
- two dice: 50
- three dice: 85
- four dice: 140
- five dice: 200
- six dice: 400

## Chance of farkling, for each # of dice rolls: 

```{r}
dice = 1:6

for (d in dice) {
  nsim = 10000
  nfarkle = 0
  for (i in 1:nsim) {
    rolls = sample(6, d, replace = T)
    sc = score(rolls)
    if (sc == 0) {
      nfarkle = nfarkle + 1
    }
  }
  print(paste(d, 'dice roll has average farkle rate:', nfarkle / nsim))
}
```

Clearly the superior method would be to derive everything, but that's a problem for tomorrow Leah. 

